PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd

#Clean Age variable
Age outlier
```{r age outlier}
# define age limits 
age_upper_limit = 117 # oldest person in the world is 117 years
age_lower_limit = 16

@@

# Inspect colnames
colnames(dem.raw.id)
```

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category)

}
```

>>>>

# RAMP demographics
## Read in data
```{r ramp dem cols}
if(RAMP == TRUE){
  ramp.dem.raw <- readRDS(file = paste0(data.raw_path, "/ramp/dem_ramp.rds"))
  dim(ramp.dem.raw)
  colnames(ramp.dem.raw)
}
```

```{r}

# Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.ramp")

ramp.dem.raw$externalDataReference <- ramp.dem.raw$`Login ID` #move Login ID to external data reference

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.ramp, #Date when participant signed up to ramp
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
     # EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
     # demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
     # demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
    #  demographics.o_levelsgcses_or_equivalent, # level of education - gcse
     # demographics.cses_or_equivalent, # level of education - cse
     # demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.dem.raw.id)
  # Inspect colnames
  colnames(ramp.dem.raw.id)
  #Differences
  dim(ramp.dem.raw)[1]-dim(ramp.dem.raw.id)[1]
}
```

```{r ramp copy dem}
if(RAMP == TRUE){
  dem.raw.id <- ramp.dem.raw.id
}
```

#Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# define age limits 
age_upper_limit = 117 # oldest person in the world is 117 years
age_lower_limit = 16

@@

# Inspect colnames
colnames(dem.raw.id)
}
```

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )
}

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category_full =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category_full)

# RAMP employment
## Read in data
```{r ramp emp cols}
if(RAMP == TRUE){
  ramp.employment.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
  dim(ramp.employment.raw)
  colnames(ramp.employment.raw)
}
```

# Select variables
```{r ramp employment cleaning}
if(RAMP == TRUE){
  exclude_cols_employment <- c("ID",
                        "Sample",
                        "startDate.ramp")

ramp.employment.raw$externalDataReference <- ramp.employment.raw$`Login ID` #move Login ID to external data reference

ramp.employment.raw.id <- ramp.employment.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample,
      ID,
      startDate.ramp,
      highest_education_ramp = employment.what_is_your_highest_level_of_education, #this is missing NVQ or equivalent which we have in GLAD-COPING
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_employment) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

```{r ramp copy emp}
if(RAMP == TRUE){
  employment.raw.id <- ramp.employment.raw.id
}
```
